Globally-centered autocovariances in MCMC

### 2020-08-17 Paper/main.tex
Pushing current version

\usepackage{amssymb,amsthm,amsmath}
\usepackage{commath,hyperref}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
\usepackage[linesnumbered, ruled]{algorithm2e}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.
\allowdisplaybreaks
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

@@

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu_g$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Upsilon(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k$. For the $s$th chain, it is given by
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}
%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_ACvF} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

@@

A critical use of autocovariances is in the assessment of Monte Carlo variability of estimators.  Let $g:\X \longrightarrow \mathbb{R}^p$ be an $F$-integrable function so that interest is in estimating
%
\[
\mu_g = \mathbb{E}_F[g(X)] = \int_{\X}g(x)F(dx)\;.
\]

Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$. Let $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1.

Reliability of $\bar{\bar{Y}}$ is assessed by its Monte Carlo variability via an asymptotic sampling distribution \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. 
% To assess the reliability of simulations, we are interested in Monte Carlo error, i.e. $\bar{Y_s} - \mu$ for the $s^{th}$ chain. 
% The sampling distribution for Monte Carlo error is available via Markov chain CLT (\cite{jones2004markov})
 A Markov chain central limit theorem (CLT) holds if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu_g) \xrightarrow{d} N(0,\Sigma)\,
\end{equation}
where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right) := \sum_{k = -\infty}^{\infty}\Upsilon (k)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$. There is a rich literature on estimators of $\Sigma$ available for single-chain MCMC implementations. The most common are spectral variance estimators (SVE) \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from multiple parallel chains. Batch means estimators are computationally more efficient than SVE, whereas SVE estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

SV estimators are constructed as weighted and truncated sums of estimated lag covariances. The locally and globally centered empirical estimators for $\Upsilon(k)$, $\hat{\Upsilon}_s(k)$ and $\hat{\Upsilon}_{G,s}(k)$, can be defined in a similar way as $\hat{\Gamma}_s$ and $\hat{\Gamma}_{G,s}$ in Equation~\ref{eq:empirical_ACvF},\ref{eq:G-ACvF} for the Markov chain $\{Y_{st}\}_{t \geq 1}$ for all $s$. Let $w: \mathbb{R} \to [-c,c]$ be a lag window function and $b_n \in \mathbb{N}$ be a truncation point. The SV estimator of $\Sigma$ from a single chain is
%

\begin{equation} \label{eq:sve}
    \hat{\Sigma}_{s} = \sum_{k=-n+1}^{n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_s(k)
\end{equation}

An overall estimate of $\Sigma$ from all $m$ chains is the average SV estimator (ASV):
\[
\hat{\Sigma}_A = \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{s}\;.
\]
%  We propose a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

% Recall from Equation \ref{eq:empirical_ACvF} that $\hat{\Gamma}(k)$ is the empirical autocovariance function given by

% \begin{equation}
%     \hat{\Gamma}(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}}(X_t - \bar{X})(X_{t + \abs{k}} - \bar{X})^T
% \end{equation}

We define the globally-centered SV estimator as the weighted and truncated sum of G-ACvFs
%
% Consider $m$ Markov chains where $\hat{\Gamma}_s(k)$, $\hat{\Gamma}_{G,s}(k)$ and $\hat{\Sigma}_{S, s}$ are the lag-$k$ empirical ACF, G-ACF, and SVE respectively for $s^{th}$ Markov chain, $s\in {1,...,m}$. 
% Consider the globally-centered SV estimator (G-SVE) for the $s$th chain using G-ACF:
%
\begin{align*}
    \hat{\Sigma}_{G} &= \sum_{k=-b_n+1}^{b_n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_{G}(k)\,.
\end{align*}
% and the combined G-SVE is
% \[
% \hat{\Sigma}_{G} =  \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{G,s}\,.
% \]
For $\hat{\Sigma}_{G}$ to accurately estimate $\Sigma$, conditions need to be imposed on the weight function, $w$. 
%

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is a continuous at $0$ and at all but finite number of points. It is bounded and even function with $w(0)=1, \; \abs{w(x)} < c$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W = \sum_{k = -\infty}^{\infty} w(k/b_n)$ is finite.
\end{ass}
% We will consider a range of lag windows $w(x)$; where $w(x)$ is a continuous, even function with $w(0)=1, \abs{w(x)} < c$, $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W_n < \infty$. 
Assumption~\ref{ass:lag_window} is standard \citep[see]{ande:1971} and is satisfied by most lag windows. We will use the popular Bartlett lag window in our simulations. The function $w(.)$ is given by
\[
w(x) = \begin{cases}
\left(1 - \abs{x}\right) & \textrm{for } \abs{x} \leq 1 \\
0 & \textrm{otherwise}\,.
\end{cases}
\]

In relation to lag window, we are interested in $w(k/b_n)$ for lag-$k$. Therefore, zero weightage is given to autocovariances for lag greater than the truncation point $b_n$. This is true for Bartlett but not true for the quadratic spectral lag window in \cite{andr:1991} for example.

@@

We will present three main results for the G-SV estimator. First, we provide conditions for strong consistency. Strong consistency is particularly important to demonstrate for MCMC as it is required to ensure sequential stopping rules yield correct coverage at termination \citep{glyn:whit:1992}. A critical assumption is that of a strong invariance principle below. Let $B(n)$ denotes a standard $p$-dimensional Brownian motion.

@@

\begin{theorem}
\label{th:consistency}
 Let the assumptions of Theorem~\ref{thm:kuelbs} hold with $q = 1$. If $\hat{\Sigma}_{SV,s} \xrightarrow{a.s.} \Sigma$ for all $s$, and $n^{-1}{b_n \log \log n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{G} \to \Sigma$ with probability 1, as $n \to \infty$.
\end{theorem}

% \begin{proof}
% We complete the proof by showing that $\hat{\Sigma}_{G}^{ij} - \tilde{\Sigma}^{ij}$ converges to 0 almost surely for all $i,j \in \{1,...,p\}$ and $\tilde{\Sigma}^{ij} \to \Sigma^{ij}$ with probability 1 as $n \to \infty$. We will show that

% \[
% \hat{\Sigma}_{G-SVE}^{ij} - \tilde{\Sigma}^{ij} = g_1(n)D^2 + g_2(n)D + g_3(n)
% \]
% where $D$ is the finite random variable associated with SIP. We will show that
% \begin{align*}
%     g_1(n) &= (4+C_1)\dfrac{b_n \psi^2(n)}{n^2} - 4\dfrac{\psi^2(n)}{n^2}\\
%     g_2(n) &= 2\sqrt{2}\|L\|p^{1/2}(1+\epsilon)\left[(4+C_1)\dfrac{b_n\psi(n)\sqrt{n\log \log n}}{n^2} - 4\dfrac{\psi(n)\sqrt{n\log \log n}}{n^2}\right]\\
%     g_3(n) &= \|L\|^2 p (1+\epsilon)^2\left[(4+C_1)\dfrac{b_n \log\log n}{n} - 4 \dfrac{\log \log n}{n}\right]
% \end{align*}

% Using $\psi(n) = \mathcal{O}(\sqrt{n \log \log n}),\; b_n = o(n), \textrm{ and } b_n \log \log n = o(n)$, we can finally we prove that $g_1(n), g_2(n), g_3(n) \to 0$ with probability 1, as $n \to \infty$. The proof is in Appendix subsection \ref{appendix:strong_consis}
% \end{proof}

% We are interested in the finding the bias for G-SVE estimator as a function of $n$. 
%\begin{ass} \label{ass:bias}
    % Let $\Gamma_s(k)$ be the lag-k autocovariance for $s^{th}$ Markov chain and $w(x)$ be the lag window. We assume that there exists a $q \geq 0$ such that for all $s$
    % \begin{enumerate} [a.]
    %     \item $\sum\limits_{-\infty}^{\infty}\abs{k}^q \|\Gamma_s(k)\| < \infty$ {\color{blue} DV: This assumption is satisfied by the polynomial ergodicity, so can remove it.}
    %     \item $\lim _{x \to 0}\dfrac{1 - w(x)}{\abs{x}^q} = k_q < \infty$ 
%        \item $\dfrac{b_n^q}{n} \to 0$ as $n \to \infty$
%   \end{enumerate}

    % If $k_q$ is finite for some $q_o$, then it is zero for $q < q_0$. $q$ is taken to be the largest number satisfying the first two conditions above.
%\end{ass}

\begin{theorem}\label{th:G-SVE_bias}
Let the assumptions of Theorem~\ref{thm:kuelbs} hold with $q$ such that $\lim _{x \to 0}{1 - w(x)}/\abs{x}^q = k_q < \infty$ and $b_n^q/n \to 0$ as $n \to \infty$. Then
\[
 \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] = -k_q\Phi^{(q)}\,.
 \]
\end{theorem}

\begin{theorem} \label{th:G-SVE_variance}
 Let the assumptions of Theorem~\ref{thm:kuelbs} hold and let $\E[D^4] < \infty$  and  $\E \|Y_{11}\|^4 < \infty$, then $b_n^{-1}{n}\Var \left(\hat{\Sigma}_{G}^{ij} \right) = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx  + o(1)$.
\end{theorem}

% \begin{proof}
% Recall that $\tilde{\Sigma}$ is the averaged pseudo spectral variance estimator which uses the unobserved mean. Further, let $\tilde{\Sigma}^{ij}$ be the $(i,j)^{th}$ element of this matrix. We will prove that the variance of $\hat{\Sigma}_{G} \textrm{ and } \tilde{\Sigma}$ are equivalent for large sample sizes. The proof is in Appendix subsection \ref{appendix:variance}
% \end{proof}

@@

The SV estimator, despite good statistical properties, poses application limitations due to slow computation. From equation \ref{eq:empirical_ACvF} and \ref{eq:sve}, the computation of the SV estimator has a complexity of $\mathcal{O}(n^2 p^2)$ in general, and for Bartlett (and other truncated lag windows) $\mathcal{O}(b_n n p^2)$. In MCMC typically, $n$ is large, for highly-correlated processes, $b
_n$ is large, and for high-dimensional target distributions, $p$ could also be large.

To overcome this limitation, we adapt the  fast Fourier transform based algorithm of \cite{heberle2017fast} for our purposes. Suppose $w_k = w(k/b_n)$ and let $T(w)$ be the $n \times n$ Toeplitz matrix of weights with the first column being $(1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1})^T$. Notice an alternate formulation of SVE for the $s^{th}$ chain $\{Y_{st};\, t \in 1, ..., n\}$ i
%
\begin{equation} \label{eq:kyriakoulis}
    \hat{\Sigma}_s = \dfrac{1}{n}A_s^T T(w) A_s, \qquad \textrm{ where } A_s = \begin{pmatrix}
    Y_{s1} - \bar{Y}_s  & \dots & Y_{sn} - \bar{Y}_s
\end{pmatrix}^T
\end{equation}

>>>>

\usepackage{amssymb,amsthm,amsmath}
\usepackage{commath,hyperref}
\usepackage{enumerate,xcolor}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{caption,setspace}
\usepackage{subcaption}
\DeclareMathOperator{\Tr}{Tr}
\newcommand{\dd}[1]{\mathrm{d}#1}
\newcommand{\ds}{\displaystyle}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Var}{\text{Var}}
\newcommand{\Cov}{\text{Cov}}
\newcommand{\X}{\mathcal{X}}
\newcommand{\B}{\mathcal{B}}
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}
% These three packages are from the American Mathematical Society and includes all of the important symbols and operations 
\usepackage{fullpage}
\usepackage{comment}
\usepackage[linesnumbered, ruled]{algorithm2e}
% By default, an article has some vary large margins to fit the smaller page format.  This allows us to use more standard margins.
\allowdisplaybreaks
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}

@@

Let $F$ be the target distribution with mean $\mu$ defined on the set $\X \subseteq \mathbb{R}^p$ equipped with a countably generated Borel $\sigma$-field $\mathcal{B}(\X)$. For $s = 1, \dots, m$, let $\{X_{st}; \; t \in \mathbb{Z}\}$ be the $s^{th}$ $F$-Harris ergodic stationary Markov chain \citep[see][for definitions]{meyn:twee:2009} employed to learn characteristics about $F$.
% If $\bar{X} = n^{-1} \sum_{t=1}^{n}X_t$ , then with probability 1, $\bar{X} \to \mu_g$ as $n \to \infty$. \\
The process is covariance stationary so that
the ACvF at lag $k$ depends only on $k$ and is defined as 
%
\[
    \Gamma(k) = \Cov_F(X_{s1}, X_{s\,1+k})= \mathbb{E}_F \left[(X_{s1} - \mu)(X_{s\,1+k} - \mu)^T \right]\,.
\]
Estimating $\Gamma(k)$ is critical to assessing the quality of the sampler and the quality of sample statistics. Let $\bar{X}_s = n^{-1} \sum_{t=1}^{n} X_{st}$ denote the Monte Carlo estimator of $\mu$ from the $s$th chain. The standard estimator for $\Gamma(k)$ is the sample autocovariance matrix at lag $k$. For the $s$th chain, it is given by
%
\begin{equation} \label{eq:empirical_ACvF}
    \hat{\Gamma}_s(k) = \dfrac{1}{n}\sum_{t=1}^{n-\abs{k}} \left(X_{st} - \bar{X}_s \right) \left(X_{s\, t+\abs{k}} - \bar{X}_s \right)^T\,.
\end{equation}
{\color{red} We need to figure out how we want to define autocovariance: with or without absolute value.}
%
% It is a known fact that most of the Markov chains obtained from practical MCMC algorithms are not stationary. Regardless, Equation~\ref{eq:empirical_ACvF} provides the empirical estimator for ACF at a particular lag for the Markov chain with any initial distribution (\cite{geyer2011introduction}).\\

@@

A critical use of autocovariances is in the assessment of Monte Carlo variability of estimators.  Let $g:\X \longrightarrow \mathbb{R}^p$ be an $F$-integrable function so that interest is in estimating
$\mu_g = \mathbb{E}_F[g(X)]$.
% \[
% \mu_g = \mathbb{E}_F[g(X)] = \int_{\X}g(x)F(dx)\;.
% \]
 % 
Set $\{Y_{st}\}_{t \geq 1} = \{g(X_{st})\}_{t \geq 1}$ for all $s = 1, \dots, m$. Let $\bar{Y}_s = n^{-1}\sum_{t=1}^{n}Y_{st}$  and  $\bar{\bar{Y}} = m^{-1}\sum_{s=1}^{m}\bar{Y}_s$. By Birkhoff's ergodic theorem,  $\bar{\bar{Y}} \to \mu_g$ as $n \to \infty$ with probability 1.

An asymptotic sampling distribution may be available via a Markov chain central limit theorem (CLT) if there exists a $p \times p$ positive-definite matrix $\Sigma$ such that
%
\begin{equation}
\label{eq:CLT}
  \sqrt{mn}(\bar{\bar{Y}} - \mu_g) \xrightarrow{d} N(0,\Sigma)\,
\end{equation}
where
\begin{equation}
\label{eq:sigma}
  \Sigma = \sum_{k = -\infty}^{\infty} \Cov_F \left( Y_{11}, Y_{1(1+k)} \right) := \sum_{k = -\infty}^{\infty}\Upsilon (k)
\end{equation}
%
The goal in output analysis for MCMC is to estimate $\Sigma$ in order to assess variability in $\bar{\bar{Y}}$ \citep{fleg:hara:jone:2008,roy:2019,vats:rob:fle:jon:2020}. There is a rich literature on estimating $\Sigma$ for single-chain MCMC implementations. The most common are spectral variance (SV) estimators \citep{andr:1991,vats:fleg:jon:2018} and batch means estimators \citep{chen:seila:1987,vats:fleg:jon:2019}. Recently, \cite{gupta:vats:2020} constructed a replicated batch means estimator for estimating $\Sigma$ from parallel chains. Batch means estimators are computationally more efficient than SV estimators, whereas SV estimators are more reliable \citep{fleg:jone:2010}. Here, we utilize G-ACvF estimators to construct globally-centered SV estimator (G-SVE) of $\Sigma$. Using the methods of \cite{heberle2017fast}, we also provide an efficient algorithm for obtaining the G-SV estimator.

The locally and globally centered  estimators for $\Upsilon(k)$ are $\hat{\Upsilon}_s(k)$ and $\hat{\Upsilon}_{G,s}(k)$, respectively, where
\[
\hat{\Upsilon}_s(k) = \dfrac{1}{n} \ds \sum_{t=1}^{n - |k|} (Y_{st} - \bar{Y}_s)(Y_{s\,t+k} - \bar{Y}_s)^T \quad \text{ and } \quad \hat{\Upsilon}_{G,s}(k) = \dfrac{1}{n} \ds \sum_{t=1}^{n - |k|} (Y_{st} - \bar{\bar{Y}})(Y_{s\,t+k} - \bar{\bar{Y}})^T\,,
\] 
and let $\hat{\Upsilon}_G(k) = m^{-1} \sum_{k=1}^{m}\hat{\Upsilon}_{G,s}(k)$. SV estimators are constructed as weighted and truncated sums of estimated lag covariances. For some $c \geq 1$, let $w: \mathbb{R} \to [-c,c]$ be a lag window function and $b_n \in \mathbb{N}$ be a bandwidth. The SV estimator of $\Sigma$ for chain $s$ is
%
%
\begin{equation} \label{eq:sve}
    \hat{\Sigma}_{s} = \sum_{k=-n+1}^{n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_s(k)\,.
\end{equation}
Large-sample properties of $\hat{\Sigma}_s$ have been studied widely. \cite{vats:fleg:jon:2018} provided conditions for strong consistency while \cite{fleg:jone:2010,vats:flegal:2018} provided large-sample bias and variance results. These results also extend naturally to an average SV (ASV) estimator
\[
\hat{\Sigma}_A = \dfrac{1}{m}\sum_{s=1}^{m}\hat{\Sigma}_{s}\;.
\]

%  We propose a globally-centered spectral variance estimator (G-SVE) wherein we use G-ACF to estimate the autocovariance. Strong consistency for MSVE has been addressed by \cite{vats:fleg:jon:2018}. Bias and variance calculations for a variant of SVE with known mean is done by \cite{hannan2009multiple}. In this paper, we will address the asymptotic properties (strong consistency), bias and variance calculations for G-SVE.

@@

We define the G-SV estimator as the weighted and truncated sum of G-ACvFs
%
\begin{align*}
    \hat{\Sigma}_{G} &= \sum_{k=-n+1}^{n-1}w\left(\dfrac{k}{b_n}\right)\hat{\Upsilon}_{G}(k)\,.
\end{align*}
%
To accurately estimate $\Sigma$, standard conditions are imposed on the lag window, $w$. 
%

\begin{ass}
\label{ass:lag_window}
The lag window $w(x)$ is continuous at $0$ at all but finite number of points. Further, $w(x)$ is a bounded and even function with $w(0)=1, \; {\color{red}\abs{w(x)} < c}$, \; $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W = \sum_{k = -\infty}^{\infty} w(k/b_n)$ is finite.
\end{ass}
% We will consider a range of lag windows $w(x)$; where $w(x)$ is a continuous, even function with $w(0)=1, \abs{w(x)} < c$, $\int_{-\infty}^{\infty}w^2(x)dx < \infty$, and $W_n < \infty$. 
Assumption~\ref{ass:lag_window} is standard \citep[see]{ande:1971} and is satisfied by most lag windows. We will use the following popular Bartlett lag window in our simulations:
\[
w(x) = \begin{cases}
1 - \abs{x} & \textrm{for } \abs{x} \leq 1 \\
0 & \textrm{otherwise}\,.
\end{cases}
\]

% In relation to lag window, we are interested in $w(k/b_n)$ for lag-$k$. Therefore, zero weightage is given to autocovariances for lag greater than the bandwidth $b_n$. This is true for Bartlett but not true for the quadratic spectral lag window in \cite{andr:1991} for example.

@@

We will present three main results for the G-SV estimator. First, we provide conditions for strong consistency. Strong consistency is particularly important to ensure sequential stopping rules yield correct coverage at termination \citep{glyn:whit:1992}. A critical assumption is that of a strong invariance principle which the following theorem establishes. Let $B(n)$ denotes a standard $p$-dimensional Brownian motion.

@@

\begin{theorem}
\label{th:consistency}
 Let the assumptions of Theorem~\ref{thm:kuelbs} hold with $q = 1$. If $\hat{\Sigma}_{s} \xrightarrow{a.s.} \Sigma$ for all $s$, and $n^{-1}{b_n \log \log n} \to 0 \textrm{ as } n \to \infty$, then $\hat{\Sigma}_{G} \overset{a.s.}{\to} \Sigma$ as $n \to \infty$.
\end{theorem}

Our next two results establish large-sample bias and variance for G-SV and mimic those of $\hat{\Sigma}_s$ which can be found in \cite{hannan:1970}. However, \cite{hannan:1970} proved the results assuming $\mu_g$ was known. Here, we present the results under mild conditions for when $\mu_g$ is replaced by $\bar{\bar{Y}}$. Let $\Sigma^{ij}$ and $\hat{\Sigma}_G^{ij}$ denote the $ij$th element of the matrix $\Sigma$ and $\hat{\Sigma}_G$, repectively.

\begin{theorem}\label{th:G-SVE_bias}
Let the assumptions of Theorem~\ref{thm:kuelbs} hold with $q$ such that 
\[
\lim _{x \to 0} \dfrac{1 - w(x)}{\abs{x}^q} = k_q < \infty
\] 
and $b_n^q/n \to 0$ as $n \to \infty$. Then $ \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] = -k_q\Phi^{(q)}\,.$
% \[
%  \lim_{n \to \infty}b_n^q\mathbb{E} \left[\hat{\Sigma}_{G} - \Sigma \right] = -k_q\Phi^{(q)}\,.
%  \]
\end{theorem}

\begin{theorem} \label{th:G-SVE_variance}
 Let the assumptions of Theorem~\ref{thm:kuelbs} hold and let $\E[D^4] < \infty$  and  $\E \|Y_{11}\|^4 < \infty$, then $ \lim_{n \to \infty} b_n^{-1}{n}\Var \left(\hat{\Sigma}_{G}^{ij} \right) = [\Sigma_{ii}\Sigma_{jj} + \Sigma_{ij}^2]\int_{-\infty}^{\infty}w(x)^2dx $.
\end{theorem}

@@

The SV estimator, despite good statistical properties, poses application limitations due to slow computation. From \eqref{eq:empirical_ACvF} and \eqref{eq:sve}, the complexity of the SV estimator is $\mathcal{O}(n^2 p^2)$ in general, and for Bartlett (and other truncated) lag window, the complexity is $\mathcal{O}(b_n n p^2)$. In MCMC, and especially for slow mixing Markov chains, $n$ and $b
_n$ are large, limiting the use of SV estimators.

To overcome this, we adapt the  fast Fourier transform based algorithm of \cite{heberle2017fast} for our purposes. Suppose $w_k = w(k/b_n)$ and let $T(w)$ be the $n \times n$ Toeplitz matrix of weights with the first column being $(1 ~ w_1 ~ w_2 ~ \dots, ~ w_{n-1})^T$. Notice an alternate formulation of $\hat{\Sigma}_s$
%
\begin{equation} \label{eq:kyriakoulis}
    \hat{\Sigma}_s = \dfrac{1}{n}A_s^T T(w) A_s, \qquad \textrm{ where } A_s = \begin{pmatrix}
    Y_{s1} - \bar{Y}_s  & \dots & Y_{sn} - \bar{Y}_s
\end{pmatrix}^T
\end{equation}
